Optimizing Pypy's Garbage Collector for Copy On Write Performance given a forky Web-Server Workload

### 2018-03-23 paper/main.tex
Final draft of paper

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

@@

This paper examines an optimization in PyPy's Generational Garbage Collector\cite{pypy-doc} which is designed to reduce the number of shared pages that get disturbed during a major collection.  To accomplish this, we take the Garbage Collector state which was previously stored within each object and we move that state into a dense index stored in its own page in memory.  This prevents the Garbage Collector from needing to write to shared memory during collection.

@@

\begin{lstlisting}[language=python]
def obj_visited(self, obj):
  return (
    self.visited_objs.contains(obj) and
    self.visited_objs.get(obj) == obj)

@@

def unmark_obj_visited(self, obj):
  self.visited_objs.setitem(obj,
        llmemory.NULL)
\end{lstlisting}

@@

\begin{lstlisting}[language=python]
class LargeObject():
  # Use a bunch of memory
  def __init__(self):
    self.lists = []
    self.strs = []
    for i in range(1600):
      self.lists.append(['text'] * i)
      for j in range(40):
        self.strs.append('str' * 8)

@@

def on_page_load():
  global glob_a, glob_b, glob_c
  glob_a, glob_b, glob_c = (
    LargeObject(), glob_a, glob_b)
\end{lstlisting}

@@

To track memory usage as the number of requests increases, we make a burst of 100 requests and then check the memory usage of each uWSGI worker process.  We repeat this process until we've made 50 bursts of requests or 5000 requests in total.  At each step we check memory usage by scraping that process's $/proc/PID/smaps$ file.  The private memory used by that process is the sum of all lines starting with ``Private\_clean'' and ``Private\_dirty'' and the public memory used by that process is the sum of all lines starting with ``Public\_clean'' and ``Public\_dirty''.

>>>>

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codered}{rgb}{0.6,0,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

% Python style for highlighting
\lstdefinestyle{pystyle}{
    language=python,
    otherkeywords={self},
    keywordstyle=\color{codered},
    emphstyle=\color{codepurple},    % Custom highlighting style
    stringstyle=\color{codegreen},
    frame=tb,                         % Any extra options here
    showstringspaces=false            % 
}

\lstset{style=pystyle}

@@

This paper examines an optimization in PyPy's Generational Garbage Collector\cite{pypy-doc} which is designed to reduce the number of shared pages that get disturbed during a major collection.  To accomplish this, we take the Garbage Collector state which was previously stored within each object and we move that state into a dense index stored in its own page in memory.  This prevents the Garbage Collector from needing to write to shared memory during collection.  We then provide experimental evidence showing that this optimization does in fact reduce memory usage of a group of processes.

@@

\noindent
\begin{minipage}{\linewidth}
\begin{lstlisting}[language=python]
def obj_visited(self, obj):
  return (
    self.visited_objs.contains(obj) and
    self.visited_objs.get(obj) == obj)

@@

def unmark_obj_visited(self, obj):
  self.visited_objs.setitem(obj,
        llmemory.NULL)
\end{lstlisting}
\end{minipage}

@@

\noindent
\begin{minipage}{\linewidth}
\begin{lstlisting}[language=python]
class LargeObject():
  # Use a bunch of memory
  def __init__(self):
    self.lists = []
    self.strs = []
    for i in range(1600):
      self.lists.append(["text"] * i)
      for j in range(40):
        self.strs.append("str" * 8)

@@

def on_page_load():
  global glob_a, glob_b, glob_c
  glob_a, glob_b, glob_c = (
    LargeObject(), glob_a, glob_b)
\end{lstlisting}
\end{minipage}

@@

To track memory usage as the number of requests increases, we make a burst of 100 requests and then check the memory usage of each uWSGI worker process.  We repeat this process until we've made 50 bursts of requests or 5000 requests in total.  At each step we check memory usage by scraping that process's $/proc/PID/smaps$ file.  The private memory used by that process is the sum of all lines starting with ``Private\_clean'' and ``Private\_dirty'' and the shared memory used by that process is the sum of all lines starting with ``Shared\_clean'' and ``Shared\_dirty''.
